### 2011-04-05 docs/enthalpy/enthalpy.tex
Fix sign error in (12)
More comments by Ann.

If we set $\omegadot_k = \Hnuc = 0$ for simplicity, then the constraint as derived using $h$ can be written 
\begin{equation}
\nabla \cdot \Ub
- \left( \frac{(1 - \rho h_p )p_T - \rho c_p}{\rho^2
  c_p p_\rho} \right) \frac{D p_0}{D t} = 0
\end{equation}
and the constraint derived using $e$ can be written
\begin{equation}
\nabla \cdot \Ub 
+ \left( \frac{\rho e_T + \rho e_p p_T}{\rho^2 e_T p_\rho + p p_T} \right) \frac{D p_0}{D t} = 0
\end{equation}

@@

The mismatch between the pressure implicit in the definition of $h$
and $p_0$ can be seen by substituting $h = e + p/\rho$ into the
enthalpy equation, where we replace $p$ with $p_0$ in the $Dp/Dt$ term:
\begin{eqnarray}
\frac{\partial(\rho h)}{\partial t} &=& -\nabla\cdot(\rho h\Ub) + 
  \frac{Dp_0}{Dt} + \rho\Hnuc \nonumber \\
%
\frac{\partial(\rho e)}{\partial t} + \frac{\partial p}{\partial t} &=&
 -\nabla\cdot(\rho e\Ub) -\nabla\cdot(p\Ub) + \frac{Dp_0}{Dt} + \rho\Hnuc \nonumber \\
%
\frac{\partial(\rho e)}{\partial t} &=&
 -\nabla\cdot(\rho e\Ub) - p\nabla\cdot\Ub + \rho\Hnuc + 
  \left \{ \frac{Dp_0}{Dt} - \frac{Dp}{Dt} \right \} \nonumber 
\end{eqnarray}
We note that the last equation appears to be an internal energy
evolution equation, but with an additional forcing term: $D(p_0 - p)/Dt$.

@@

\item We substitute $p$ for $p_0$ in two places: 1) in the derivation of the
constraint, 2) in the enthalpy evolution equation.   Are these substitutions
consistent with each other?  If they are not consistent, which of these 
substitutions is responsible for filtering the acoustics from the system?

\item In the EOS, we take $h$ and construct $e$ by subtracting off
  $p$---depending on the inputs to the EOS, we treat the `$p$' part of $h$
  differently.

   if we come into the EOS with $\rho$, $T$, then we get $p$ and $e$ from
   the EOS and compute $h$.

   if we come into the EOS with $\rho$, $h$, then we find a $T$ such that
   $e + p(\rho,T)/\rho = h$, where $p(\rho,T)$ is the pressure from the EOS.
   It is not necessarily equal to $p_0$.

   if we come into the EOS with $\rho$, $p_0$, then find a $T$ that matches
   $p_0$, and then $h = e + p_0/\rho$.  In this case, there is no mismatch
   in the enthalpy equation.

\item do we want to always insist that the `$p$' hidden in the definition of
   $h$ is $p_0$?

>>>>

If we set $\omegadot_k = \Hnuc = 0$ for simplicity, then the constraint as derived using $h$ can be written 
\begin{equation}
\nabla \cdot \Ub
+ \left( \frac{(1 - \rho h_p )p_T - \rho c_p}{\rho^2
  c_p p_\rho} \right) \frac{D p_0}{D t} = 0
\end{equation}
and the constraint derived using $e$ can be written
\begin{equation}
\nabla \cdot \Ub 
+ \left( \frac{\rho e_T + \rho e_p p_T}{\rho^2 e_T p_\rho + p p_T} \right) \frac{D p_0}{D t} = 0
\end{equation}

We note that if we evaluate both constraints for $p = \rho R T,$ with $h = c_p T,$ $e = c_v T,$
$c_p = c_v + R$ and $\gamma = c_p / c_v,$ then both constraints reduce to 
\begin{equation}
\nabla \cdot \Ub + \frac{1}{\gamma p} \frac{D p_0}{D t} = 0
\end{equation}

@@

The mismatch between the pressure implicit in the definition of $h$
and $p_0$ can be seen by substituting $h = e + p/\rho$ into the
enthalpy equation, where we replace $p$ with $p_0$ in the $Dp/Dt$ term:
\begin{eqnarray}
\frac{\partial(\rho h)}{\partial t} &=& -\nabla\cdot(\rho h\Ub) + 
  \frac{Dp_0}{Dt} + \rho\Hnuc \nonumber \\
%
\frac{\partial(\rho e)}{\partial t} + \frac{\partial p}{\partial t} &=&
 -\nabla\cdot(\rho e\Ub) -\nabla\cdot(p\Ub) + \frac{Dp_0}{Dt} + \rho\Hnuc \nonumber \\
%
\frac{\partial(\rho e)}{\partial t} &=&
 -\nabla\cdot(\rho e\Ub) - p\nabla\cdot\Ub + \rho\Hnuc + 
  \left \{ \frac{Dp_0}{Dt} - \frac{Dp}{Dt} \right \} \nonumber 
\end{eqnarray}
However, if we solve the evolution equation for $e$ we would substitute 
$p_0$ for $p$ in this equation as well.  Thus, we can pose the situation as the following.  
If we solve the evolution equation for $h$ then we effectively are solving
\begin{equation}
\frac{\partial(\rho e)}{\partial t} +
  \nabla\cdot(\rho e\Ub) = p \; \nabla\cdot\Ub + \rho\Hnuc + 
  \left \{ \frac{Dp_0}{Dt} - \frac{Dp}{Dt} \right \} \nonumber 
\end{equation}
but if we solve the evolution equation for $e$ we are effectively solving
\begin{equation}
\frac{\partial(\rho e)}{\partial t} +
  \nabla\cdot(\rho e\Ub) = p_0 \nabla\cdot\Ub + \rho\Hnuc \nonumber 
\end{equation}

The second equation subtracted from the first gives:
\begin{equation}
  \left \{ \frac{Dp_0}{Dt} - \frac{Dp}{Dt} \right \} + (p - p_0)  \nabla \cdot \Ub
\end{equation}

Suppose we solve the current enthalpy equation, but when we call the EOS, we 
subtract $p_0$ from $\rho h$ and then call the EOS with $e$ instead of $h$.  This is equivalent
to:
\begin{eqnarray}
\frac{\partial(\rho h)}{\partial t} &=& -\nabla\cdot(\rho h\Ub) +
  \frac{Dp_0}{Dt} + \rho\Hnuc \nonumber \\
%
\frac{\partial(\rho e)}{\partial t} + \frac{\partial p_0}{\partial t} &=&
 -\nabla\cdot(\rho e\Ub) -\nabla\cdot(p_0 \Ub) + \frac{Dp_0}{Dt} + \rho\Hnuc \nonumber \\
%
\frac{\partial(\rho e)}{\partial t} &=&
 -\nabla\cdot(\rho e\Ub) - p_0 \nabla\cdot\Ub + \rho\Hnuc  \nonumber
\end{eqnarray}
which is identical to solving the energy equation.

@@

\item Should we call the EOS with $h$ as is, or call the EOS with $e = h - p_0 / \rho$?

\item When we stay on the constraint, i.e. $p_{EOS} = p_0$, then the equations for $e$ and
for $h$ are identical.   However, once we are off the constraint, do the terms in the
current evolution equation for $h$ serve to drive us back to the constraint?  Recall
our current "volume discrepancy factor" acts as a source term which modifies the
divergence constraint, which effectively modifies both $\rho$ and $T$ (or $e$ or $h$).
The term in the enthalpy equation only modifies $\rho.$  Is this relevant and/or useful??

%\item We substitute $p$ for $p_0$ in two places: 1) in the derivation of the
%constraint, 2) in the enthalpy evolution equation.   Are these substitutions
%consistent with each other?  If they are not consistent, which of these 
%substitutions is responsible for filtering the acoustics from the system?

\item Suppose we corrected the $h$ (or $e$ equation) by using the full $p$ instead
of $p_0$?  Would this be more or less consistent (one could imagine doing this as a 
correction after solving for $\pi$ earlier in the timestep).

%\item In the EOS, we take $h$ and construct $e$ by subtracting off
%  $p$---depending on the inputs to the EOS, we treat the `$p$' part of $h$
%  differently.
%
%   if we come into the EOS with $\rho$, $T$, then we get $p$ and $e$ from
%   the EOS and compute $h$.
%
%   if we come into the EOS with $\rho$, $h$, then we find a $T$ such that
%   $e + p(\rho,T)/\rho = h$, where $p(\rho,T)$ is the pressure from the EOS.
%   It is not necessarily equal to $p_0$.
%
%   if we come into the EOS with $\rho$, $p_0$, then find a $T$ that matches
%   $p_0$, and then $h = e + p_0/\rho$.  In this case, there is no mismatch
%   in the enthalpy equation.
